Apparent preferences for cognitive effort fade when multiple forms of effort

### 2021-03-15 Cost_manuscript.Rmd
new organization per experiment

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1,2^, Gary A. Kane^1,2^, & Joseph T. McGuire^1,2^"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
- \usepackage{fixltx2e}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
bibliography: cost_paper.bib
---

@@

  Cognitive and physical effort are typically regarded as costly, but recent findings have suggested that exerting effort can boost the value of prospects under certain conditions. Here we embedded mental and physical effort in a "diet choice" foraging task, which required decision makers not only to evaluate the magnitude and delay of a focal prospective reward, but also to estimate the general opportunity cost of time. In two experiments, independent sets of participants collected rewards that required equivalent periods of cognitive effort, physical effort, or unfilled delay. Monetary offers varied per trial, and the two experiments differed in whether the type of effort or delay cost was the same on every trial (between-subjects, n=21 per condition), or varied across trials (within-subjects, n=48). Participants were free either to accept or reject the cost/reward prospect offered on each trial. All participants were more likely to accept offers when rewards were higher and delays shorter (in line with a reward-maximizing strategy), almost never reversed their acceptance decisions, and committed few errors while completing the effort requirements. When participants faced only one type of cost, cognitive effort persistently produced the highest acceptance rate compared to trials with an equivalent period of either physical effort or unfilled delay. **We theorized that i**f cognitive effort were intrinsically rewarding, we would observe the same pattern of preferences when participants foraged for varying cost types in addition to rewards. In the within-subject experiment, an initially higher acceptance rate for cognitive effort trials disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions. Our results extend the view that cognitive demands may reduce the discounting effect of delays, but also suggest that differences in cost can eventually fade if individuals actively experience alternative forms of demand. Rather than assigning intrinsic value to cognitive effort, our findings support the idea that a cognitive effort requirement might influence contextual factors such as subjective delay durations. Such altered estimations can be recalibrated if multiple forms of demand are interleaved.

@@

  Evaluating whether to pursue a desirable outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards [@Shenhav2017 ;@Westbrook2015]. Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort [@Ainslie1975; @Hull1943; @Kool2010; @Kool2018; @Treadway2009; @Westbrook2013]. However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes [@HernandezLallement2014; @Kacelnik2002; @Mochon2012], and positive outcomes imbue value into their preceding effortful behavior [i.e. learned industriousness; @Eisenberger1992]. Though intuitively valid and empirically supported, this explanation leaves open questions about what features of an effortful task can make it attractive in its own right, and how contextual factors mold its perceived costs [@Inzlicht2018]. Understanding these intrinsic qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

@@

## Methods

### **Experiment 1: Between-subjects comparison of costs**

### Participants

@@

### Foraging Task

@@

include_graphics("./Images/FIG1_task_a.png")
```

### Operationalization of Cost

@@

### Analyses

@@

### **Experiment 2: Within-subjects comparison of costs**

### Participants

### Foraging Task Adaptation

### Analyses

### **Computational Modeling of Foraging Behavior**

#### **Model Fitting and Model Comparison Procedures**

## Results

### **Between-subjects: Tests of whether decision makers integrate delay and reward information**

@@

### **Between-subjects: Comparisons among the four delay and effort conditions**

@@

  We next examined the optimality of these decisions. The foraging task was configured so that there was a single reward-maximizing strategy per block type (see Operationalization of Cost in Methods), and participants were warned that accepting everything would not maximize their gains. Accordingly, participants who over- or under-accepted earned the least (Figure 3, right; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits = 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, $R^2$ = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). We performed group-level chi-squared tests of proportions to examine whether each cost biased decision makers to under- or over-accept 4 and 8 cent offers, which together should have been accepted 50% of the time. We found a significant over-acceptance of these trials by the cognitive effort group (proportion accepted = `r round(optimalProp_cost["Cognitive", "probability"], digits = 2)`%; $\chi^2$ = `r round(optimalProp_cost["Cognitive", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Cognitive", "pval"]`), but under-acceptances by the physical (proportion accepted = `r round(optimalProp_cost["Physical", "probability"], digits = 2)`%; $\chi^2$ = `r round(optimalProp_cost["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"]`) and wait (proportion accepted = `r round(optimalProp_cost["Physical", "probability"], digits = 2)`%; $\chi^2$ = `r round(optimalProp_cost["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"]`) groups (no significant difference for the Easy group). Together, these findings show that even though participants did not always perform optimally, they adequately integrated the statistics of the environment as they tracked their recent experience. Further, they suggest that cognitive effort can boost the value of an offer, as reflected by higher acceptances at the expense of earnings.

@@

### **Between-subjects: Consistency and stability**

@@

### **Within-subjects: Tests of whether decision makers integrate reward information**

@@

### **Within-subjects: Comparisons among the four delay and effort conditions**

@@

### **Consistency and stability**

@@

### **Computational modeling of foraging behavior**

``` {r OLD BTW and WTH: 16.3. OC modeling btw, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# ## fit the OC model
# summaryOC_btw <- list()
# summaryOC_btw$all <- dataBtw %>%
#                  group_by(Cost, SubjID) %>%
#                  do(optimizeOCModel(., simplify = T)) %>%
#                  ungroup()
# 
# 
# ## plot
# summaryOC_btw$plot <- ggplot(summaryOC_btw$all, aes(Cost, Gamma, fill = Cost)) +
#   # geom_hline(yintercept = 0.64, alpha = 0.9, color = "gray20", size = 1) +
#   # geom_hline(yintercept = 0.7, alpha = 0.9, color = "gray40", size = 1) +
#   # geom_hline(yintercept = 0.9, alpha = 0.9, color = "gray60", size = 1) +
#   geom_hline(yintercept = 0.74, alpha = 0.9, size = 1, linetype = "dashed") + # mean of highest earning rates across blocks
#   #geom_dotplot(stackdir = 'center', binaxis = 'y', show.legend = F) +
#   geom_boxplot(show.legend = F) +
#   geom_jitter(width = 0.1, alpha = 0.5, show.legend = F, pch = 21, size = 3) +
#   ylim(0,1.5) +
#   labs(x = "") +
#   scale_fill_manual(values = colsBtw) +
#   theme(panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.background = element_blank(),
#         axis.line = element_line(colour = "black"),
#         text = element_text(size = 16))
# 
# 
# ## one sample t against optimal strategy
# summaryOC_btw$t.test <- summaryOC_btw$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   sapply(., function(data) {t.test(data$Gamma, mu = 0.74)})
# 
# 
# ## descriptive stats
# summaryOC_btw$descriptives <- summaryOC_btw$all %>%
#   group_by(Cost) %>%
#   summarise(mGamma = round(mean(Gamma), digits = 2),
#             sdGamma = round(sd(Gamma), digits = 2))

# summaryOC$all %>%
#   ggplot(aes(Gamma, percentAccept, fill = Cost, size = temperature)) +
#   geom_point(pch = 21)

# # gamma parameter recovery tests
# # test a single subject against their own choices
# (subj <- sample(subjList_btw, 1))
# subjData <- dataBtw %>% 
#   filter(SubjID == subj) %>% 
#   group_by(Cost, Handling, Offer) %>% 
#   summarise(propAccept = mean(Choice)) %>%
#   ungroup()
# 
# params <- filter(summaryOC$all, SubjID == subj) %>% 
#   select(Gamma, temperature)
# 
# test_params(params,
#             model = expr(params$Gamma * Handling), 
#             subjData, 
#             plot = T)

# # test multiple values of gamma and plot the comparison
# lowGamma <- 0
# upperGamma <- 1.7
# nGamma <- 6
# testys <- lapply(seq(lowGamma, upperGamma, length.out = nGamma), test_params)
# (testys <- do.call(rbind, testys))
# testys %>%
#   ggplot(aes(interaction(Offer, Handling), Choice, color = Gamma)) +
#   geom_point(size = 3) +
#   geom_line(aes(group = interaction(Handling, Gamma)), size = 1) +
#   labs(x = "Offer.Handling", y = "Proportion Accepted") +
#   facet_wrap(vars(Gamma), ncol = 2) +
#   theme(legend.key = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.background = element_blank(),
#         axis.line = element_line(colour = "black"),
#         text = element_text(size = 16))

``` {r OLD BTW and WTH: 16.3. OC comparisons btw, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# # ANOVA
# OCanova <- list()
# OCanova$aov <- summary(aov(Gamma ~ Cost, data = summaryOC_btw$all))
# Fval <- round(OCanova$aov[[1]]$`F value`[1], digits = 2) # doing separate variables because I can't easily use the `` in markdown.
# Pval <- OCanova$aov[[1]]$`Pr(>F)`[1]
# Rsquared_aov <- round(OCanova$aov[[1]]$`Sum Sq`[1] / sum(OCanova$aov[[1]]$`Sum Sq`), digits = 2)
# 
# # Post-hoc pairwise tests
# OCanova$perms <- summaryOC_btw$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   do.call(cbind, .) %>%
#   select_if(is.numeric) %>%
#   rename(Physical = Physical.Gamma,
#          Cognitive = Cognitive.Gamma,
#          Wait = Wait.Gamma,
#          Easy = Easy.Gamma) %>%
#   outer(., ., FUN = Vectorize(permute), simple = T)
# diag(OCanova$perms) <- 1
# 
# OCanova$ES <- summaryOC_btw$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   do.call(cbind, .) %>%
#   select_if(is.numeric) %>%
#   rename(Physical = Physical.Gamma,
#          Cognitive = Cognitive.Gamma,
#          Wait = Wait.Gamma,
#          Easy = Easy.Gamma) %>%
#   outer(., ., FUN = Vectorize(cohenD)) %>%
#   abs()
```

``` {r OLD BTW and WTH: 16.3. corrplot btw, echo = FALSE, fig.align = "center", fig.width = 5, fig.height = 5}
# # plot all pairwise comparisons
# corrplot(OCanova$ES,
#          is.corr = F, 
#          p.mat = OCanova$perms,
#          type = "upper",
#          insig = "p-value",
#          sig.level = -1,
#          na.label = "square",
#          na.label.col = "grey",
#          method = "color", 
#          tl.col = "black", 
#          tl.cex = 0.8,
#          outline = T)
```

``` {r OLD BTW and WTH: 16.3. OC cross-validation btw, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# # Compute the OC for pre-midpoint
# summaryOC_btw$pre <- dataBtw %>%
#                  filter(Half == "Half_1") %>%
#                  group_by(Cost, SubjID) %>%
#                  do(optimizeOCModel(., simplify = T)) %>%
#                  ungroup() %>%
#                  select(SubjID, Gamma)
# 
# # let's add the gammas to the second half data and estimate prop predicted on the second half
# summaryOC_btw$predicted <- dataBtw %>%
#                  filter(Half == "Half_2") %>%
#                  group_by(Cost, SubjID) %>%
#                  left_join(summaryOC_btw$pre, by = "SubjID") %>%
#                  mutate(OC = Handling * Gamma,
#                         predicted = Offer > OC) %>% 
#                  summarise(percentPredicted = mean(predicted == Choice) * 100) %>%
#                  group_by(Cost) %>%
#                  summarise(mPredicted = round(mean(percentPredicted), digits = 2),
#                            sdPredicted = round(sd(percentPredicted), digits = 2))

# # plot
# (summaryOC_btw$predictPlot <- ggplot(summaryOC_btw$predicted, aes(Cost, percentPredicted, fill = Cost)) +
#                             labs(x = "") +
#                             geom_boxplot(show.legend = F) +
#                             scale_fill_manual(values = colsBtw) +
#                             labs(y = "Percent Predicted") +
#                             ylim(0,100) +
#                             theme(panel.grid.major = element_blank(),
#                             panel.grid.minor = element_blank(),
#                             panel.background = element_blank(),
#                             axis.line = element_line(colour = "black"),
#                             text = element_text(size = 16)))

``` {r OLD BTW and WTH: 16.3. OC Modeling wth, echo = FALSE, fig.align="center", fig.width=6, fig.height=4}
# # Here it might be best to estimate gamma for the wait trials only at first
# # once the wait OCs are estimated, then they can be modulated by effort
# # Estimate gamma per subject for each group separately
# summaryOC_wth <- list()
# summaryOC_wth$all <- dataWth %>%
#   mutate(Block = case_when(
#     Block %in% c(1, 2) ~ 1,
#     Block %in% c(3, 4) ~ 2,
#     Block %in% c(5, 6) ~ 3
#   )) %>%
#   #filter(Choice < 2) %>%
#   group_by(SubjID, Cost) %>% #group_by(SubjID, Cost, Block) %>%
#   do(optimizeOCModel(., simplify = T)) %>%
#   ungroup()
# 
# # plot
# summaryOC_wth$plot <- ggplot(summaryOC_wth$all, aes(Cost, Gamma, fill = Cost)) +
#     geom_boxplot(show.legend = F) +
#     geom_hline(yintercept = 0.64, alpha = 0.9, color = "gray20", size = 1) +
#     geom_hline(yintercept = 0.7, alpha = 0.9, color = "gray40", size = 1) +
#     geom_hline(yintercept = 0.9, alpha = 0.9, color = "gray60", size = 1) +
#     ylim(0,1.5) +
#     labs(x = "") +
#     scale_fill_manual(values = colsWth) +
#     theme(panel.grid.major = element_blank(),
#           panel.grid.minor = element_blank(),
#           panel.background = element_blank(),
#           axis.line = element_line(colour = "black"),
#           text = element_text(size = 16))

# try to replicate the group-level proportion of acceptances from 16.2
# this looks pretty much like the empirical proportion data, so the stochasticity is playing a major part in the estimation
# the estimate probability even matches actual proportions at the single subject level 
# logis <- mapply(function(g, t) {1 / (1 + exp((-1 * t) * (c(4, 8, 20) - (g * 10))))}, summaryOC_wth$all$Gamma, summaryOC_wth$all$Temperature)
# colnames(logis) <- unite(summaryOC_wth, test, SubjID, Cost)$test
# logis2 <- as.data.frame(logis) %>% 
#   gather(Cost, Probability) %>%
#   mutate(SubjID = substr(Cost, 1,3), 
#          Cost = substr(Cost, 5,20),
#          Reward = rep(c(4,8,20), nrow(.)/3)) %>%
#   group_by(Cost, Reward) %>%
#   summarize(propAccept = mean(Probability)) %>%
#   ggplot(logis2, aes(Reward, propAccept, color = Cost)) +
#     geom_point(size = 3, show.legend = T) +
#     geom_line(lwd = 1, show.legend = T) +
#     scale_color_manual(values = colsWth) +
#     scale_fill_manual(values = colsWth) +
#     #geom_errorbar(aes(ymin = propAccept - SE, ymax = propAccept + SE), width = 0.4, size = 1) +
#     scale_y_continuous(limits = c(-0.1,1.1), breaks = c(0,0.5,1)) +
#     scale_x_continuous(breaks = c(4, 8, 12), labels = c("4", "8", "20")) +
#     labs(y = "Proportion Accepted") +
#     theme(legend.key = element_blank(),
#           panel.grid.major = element_blank(),
#           panel.grid.minor = element_blank(),
#           panel.background = element_blank(),
#           axis.line = element_line(colour = "black"),
#           text = element_text(size = 16))
```

``` {r OLD BTW and WTH: 16.3. cross-v wth, echo = FALSE, fig.width=5, fig.height=4, fig.align="center"}
# # subdivide data
# dataTrain <- dataWth %>%
#    filter(Block < 5)
# 
# dataTest <- dataWth %>%
#   filter(Block > 4)
# 
# # estimate
# # per-cost, but consider the idea above
# OCs_validation <- list()
# temp <- dataTrain %>% filter(Cost == "Wait-C") %>% plyr::dlply("SubjID", identity)
# OCs_validation$`Wait-C` <- sapply(temp, function(Data) {optimizeOCModel(Data)$Gamma})
# temp <- dataTrain %>% filter(Cost == "Wait-P") %>% plyr::dlply("SubjID", identity)
# OCs_validation$`Wait-P` <- sapply(temp, function(Data) {optimizeOCModel(Data)$Gamma})
# temp <- dataTrain %>% filter(Cost == "Cognitive") %>% plyr::dlply("SubjID", identity)
# OCs_validation$Cognitive <- sapply(temp, function(Data) {optimizeOCModel(Data)$Gamma})
# temp <- dataTrain %>% filter(Cost == "Physical") %>% plyr::dlply("SubjID", identity)
# OCs_validation$Physical <- sapply(temp, function(Data) {optimizeOCModel(Data)$Gamma})
# 
# # reshape data
# temp <- bind_rows(OCs_validation) %>% 
#   mutate(SubjID = subjList_wth) %>%
#   gather(Cost, gamma, `Wait-C`:Physical) %>%
#   left_join(dataTest, ., by = c("SubjID", "Cost")) %>% # assign OCs to each subject per cost type
#   mutate(prediction = Offer > Handling * gamma) %>%
#   # plyr::dlply("SubjID", identity)
#   group_by(SubjID, Cost) %>%
#   summarize(predicted = mean(Choice == prediction))

# plot prediction %
# qplot(Cost, predicted, geom = "boxplot", data = temp) + 
#   ylim(0,1) +
#   theme_classic()
```

``` {r OLD BTW and WTH: 16.3. param compare wth, echo = FALSE}
# # repeated measures anova
# OCs_anova <- list()
# OCs_anova$aov <- summary(aov(Gamma ~ Cost + Error(SubjID / Cost), data = summaryOC_wth$all))
# 
# # table
# #pander(OCs_anova$aov)
# 
# # do all pairwise comparisons
# OCs_anova$gammas <- summaryOC_wth$all %>% 
#   reshape2::dcast(SubjID ~ Cost, value.var = "Gamma") %>%
#   select(-SubjID)
# 
# # perform a paired permutation among every pair of cost types
# OCs_anova$perms <- outer(OCs_anova$gammas, OCs_anova$gammas, FUN = Vectorize(permute), simple = T, paired = T)
# OCs_anova$perms[rbind(c(1,4),c(2,3),c(3,2),c(4,1))] <- NA
# 
# # and the effect sizes
# OCs_anova$ES <- abs(outer(OCs_anova$gammas, OCs_anova$gammas, FUN = Vectorize(DescTools::CohenD)))
# OCs_anova$ES[rbind(c(1,4),c(2,3),c(3,2),c(4,1))] <- NA
```

``` {r OLD BTW and WTH: 16.3. corrplot wth, echo = FALSE, fig.align="center", fig.width=5, fig.height=5}
# corrplot(OCs_anova$ES,
#          is.corr = F, 
#          p.mat = OCs_anova$perms,
#          type = "upper",
#          insig = "p-value",
#          sig.level = -1,
#          na.label = "square",
#          na.label.col = "grey",
#          method = "color", 
#          tl.col = "black", 
#          tl.cex = 0.8,
#          outline = T)
```

@@

include_graphics("./Images/FIG1_task_a.png")
```

@@

```{r New mixed effects plots (Fig. 4), eval = figsEnd, out.width = "80%", fig.align = "center", fig.cap = "Comparing acceptance rates among costs for each reward and handling combination. A: Proportion accepted by each group for every combination of handling time and reward. Grey dots indicate the reward-maximizing acceptance for each combination. B: Matrix portraying the coefficients (color scale) and significance that resulted from switching the reference cost condition (crosses mark non-significant comparisons). Each entry shows how much more likely a group was to accept compared to the reference condition (row). C: AIC and BIC values for a forward mixed-effects model comparison, ranging from a base intercept-only to a full-interaction model."}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG4.png")
```

@@

```{r New Prepost mixed effects plot (Fig. 6), eval = figsEnd, fig.align = "center", fig.cap = "Acceptance behavior over time in the between-subjects experiment. Left: Proportion of acceptances per handling, reward, and half, showing that effortful demands became increasingly aversive over time. Right: Mixed-effects coefficients denoting the comparison among costs acceptances for each half separately (mirrored along the diagonal). Even though acceptances steadily decreased in the effort groups, the relative preferences were preserved."}
include_graphics("./Images/FIG6.png")
```

@@

include_graphics("./Images/FIG8.png")
```

@@

```{r New Prepost mixed effects plots (Fig. 9), eval = figsEnd, fig.align = "center", fig.cap = "Acceptance rates over time in the within-subjects experiment. Top-Left: Proportion accepted in the first two and last two blocks of the experiment. An initial preference for cognitive effort trials fades amid a global decrease in acceptance rates. Top-Right: Mixed effects coefficient matrix for each half separately. The pattern from the first half resembles what was observed in the between-subjects experiment (Figure 4). Bottom: Block rder effects show attenuated differences for those who started with the physical configuration."}
include_graphics("./Images/FIG9.png")
```

@@

include_graphics("./Images/FIG10.png")

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1,2^, Gary A. Kane^1,2^, & Joseph T. McGuire^1,2^"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
- \usepackage{fixltx2e}
output:
  word_document: #default
    reference_docx: Style_reference_draft.docx
  pdf_document:
    highlight: haddock
    keep_tex: no
  html_document:
    df_print: paged
csl: apa.csl
indent: true
bibliography: cost_paper.bib
---

@@

  Cognitive and physical effort are typically regarded as costly, but recent findings have suggested that exerting effort can boost the value of prospects under certain conditions. Here we embedded mental and physical effort in a "diet choice" foraging task, which required decision makers not only to evaluate the magnitude and delay of a focal prospective reward, but also to estimate the general opportunity cost of time. In two experiments, independent sets of participants collected rewards that required equivalent periods of cognitive effort, physical effort, or unfilled delay. Monetary offers varied per trial, and the two experiments differed in whether the type of effort or delay cost was the same on every trial (between-subjects, n=21 per condition), or varied across trials (within-subjects, n=48).When participants faced only one type of cost, cognitive effort persistently produced the highest acceptance rate compared to trials with an equivalent period of either physical effort or unfilled delay. We theorized that if cognitive effort were intrinsically rewarding, we would observe the same pattern of preferences when participants foraged for varying cost types in addition to rewards. In the within-subject experiment, an initially higher acceptance rate for cognitive effort trials disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions. Our results extend the view that cognitive demands may reduce the discounting effect of delays, but also suggest that differences in cost can eventually fade if individuals actively experience alternative forms of demand. Rather than assigning intrinsic value to cognitive effort, our findings support the idea that a cognitive effort requirement might influence contextual factors such as subjective delay durations. Such altered estimations can be recalibrated if multiple forms of demand are interleaved.

@@

  Evaluating whether to pursue a desirable outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards [@Shenhav2017 ;@Westbrook2015; @Frederick2002; Walton2006]. Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort [@Ainslie1975; @Hull1943; @Kool2010; @Kool2018; @Treadway2009; @Westbrook2013]. However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes [@HernandezLallement2014; @Kacelnik2002; @Mochon2012], and positive outcomes imbue value into their preceding effortful behavior [i.e. learned industriousness; @Eisenberger1992]. Though intuitively valid and empirically supported, this explanation leaves open questions about what features of an effortful task can make it attractive in its own right, and how contextual factors mold its perceived costs [@Inzlicht2018]. Understanding these intrinsic qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

@@

## Experiment 1: Between-subjects comparison of costs

### *Methods*
#### **Participants**

@@

#### **Foraging Task**

@@

include_graphics("./Images/FIG1_task.png")
```

#### **Operationalization of Cost**

@@

#### **Analyses**

@@

### *Results*
#### **Between-subjects: Tests of whether decision makers integrate delay and reward information**

@@

#### **Between-subjects: Comparisons among the four delay and effort conditions**

@@

  We next examined the optimality of these decisions. The foraging task was configured so that there was a single reward-maximizing strategy per block type (see Operationalization of Cost in Methods), and participants were warned that accepting everything would not maximize their gains. Accordingly, participants who over- or under-accepted earned the least (Figure 3, right; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits = 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, $R^2$ = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). We performed group-level chi-squared tests of proportions to examine whether each cost biased decision makers to under- or over-accept 4 and 8 cent offers, which together should have been accepted 50% of the time. We found a significant over-acceptance of these trials by the cognitive effort group (proportion accepted = `r round(optimalProp_cost["Cognitive", "probability"], digits = 2)`; $\chi^2$ = `r round(optimalProp_cost["Cognitive", "chisquared"], digits = 2)`, *p* `r report_p(optimalProp_cost["Cognitive", "pval"])`), but under-acceptances by the physical (proportion accepted = `r round(optimalProp_cost["Physical", "probability"], digits = 2)`; $\chi^2$ = `r round(optimalProp_cost["Physical", "chisquared"], digits = 2)`, *p* `r report_p(optimalProp_cost["Physical", "pval"])`) and wait (proportion accepted = `r round(optimalProp_cost["Physical", "probability"], digits = 2)`; $\chi^2$ = `r round(optimalProp_cost["Physical", "chisquared"], digits = 2)`, *p* `r report_p(optimalProp_cost["Physical", "pval"])`) groups (no significant difference for the Easy group). Together, these findings show that even though participants did not always perform optimally, they adequately integrated the statistics of the environment as they tracked their recent experience. Further, they suggest that cognitive effort can boost the value of an offer, as reflected by higher acceptances at the expense of earnings.

@@

#### **Between-subjects: Consistency and stability**

@@

## Experiment 2: Within-subjects comparison of costs
### *Methods*
#### **Participants**

#### **Foraging Task Adaptation**

#### **Analyses**

### *Results*
#### **Within-subjects: Tests of whether decision makers integrate reward information**

@@

#### **Within-subjects: Comparisons among the four delay and effort conditions**

@@

#### **Consistency and stability**

@@

## Computational Modeling of Foraging Behavior

### *Methods*
#### **Model Fitting and Model Comparison Procedures**

### *Results*

@@

include_graphics("./Images/FIG1_task.png")
```

@@

```{r New mixed effects plots (Fig. 4), eval = figsEnd, out.width = "80%", fig.align = "center", fig.cap = "Comparing acceptance rates among costs for each reward and handling combination. A: Proportion accepted by each group for every combination of handling time and reward. Grey dots indicate the reward-maximizing acceptance for each combination. B: Matrix portraying the coefficients (color scale) and significance that resulted from switching the reference cost condition (crosses mark non-significant comparisons). Each entry shows how much more likely a group was to accept compared to the reference condition (row). C: AIC and BIC values for a forward mixed-effects model comparison, ranging from a base intercept-only to a full-interaction model."}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG4_btwlogistic.png")
```

@@

```{r New Prepost mixed effects plot (Fig. 6), eval = figsEnd, fig.align = "center", fig.cap = "Acceptance behavior over time in the between-subjects experiment. Left: Proportion of acceptances per handling, reward, and half, showing that effortful demands became increasingly aversive over time. Right: Mixed-effects coefficients denoting the comparison among costs acceptances for each half separately (mirrored along the diagonal). Even though acceptances steadily decreased in the effort groups, the relative preferences were preserved."}
include_graphics("./Images/FIG6_prepostbtw.png")
```

@@

include_graphics("./Images/FIG8_survival.png")
```

@@

```{r New Prepost mixed effects plots (Fig. 9), eval = figsEnd, fig.align = "center", fig.cap = "Acceptance rates over time in the within-subjects experiment. Top-Left: Proportion accepted in the first two and last two blocks of the experiment. An initial preference for cognitive effort trials fades amid a global decrease in acceptance rates. Top-Right: Mixed effects coefficient matrix for each half separately. The pattern from the first half resembles what was observed in the between-subjects experiment (Figure 4). Bottom: Block rder effects show attenuated differences for those who started with the physical configuration."}
include_graphics("./Images/FIG9_prepostwth.png")
```

@@

include_graphics("./Images/FIG10_modelresults.png")
